\section{Introduction}
\label{sec:intro}

In recent years, the demand for hardware with smaller form factor and higher transistor density has been steadily increasing. Consequently, it has become difficult to create high quality hardware by hand-crafting RTL designs. A behavioral synthesis tool takes a behavioral
design description (in C, C++, or SystemC), often referred to as Electronic System Level (ESL) design, and automatically generates an optimized RTL design (in hardware description languages such as VHDL or Verilog). 
Studies have shown that ESL 
reduces the design effort by $50\%$ or more while attaining
excellent performance results~\cite{Moussa99}. 
It has recently received significant attention, as the steady 
increase in hardware complexity has made it increasingly 
difficult to design high-quality designs through hand-crafted RTL under 
aggressive time-to-market schedules.  A recent example is VP9 G2 hardware
decoder IP developed by Google~\cite{googledecoder}, which has been implemented primarily in 
standard C++ and synthesized to RTL logic for different target technologies 
and performance points using Calypto's Calatpult High Level Synthesis 
tool~\cite{catapult}. Nevertheless, and in spite of availability of several
commercial behavioral synthesis tools~\cite{ctos,forte,vivado},
the adoption of the approach in main-stream hardware development for
microprocessor and SoC design companies is dependent on 
designers' confidence that the synthesized RTL indeed corresponds to the ESL
specification. 

Loop pipelining is one of the most complex transformations in behavioral synthesis. It is available in most commercial synthesis tools ({\em e.g.}, AutoESL~\cite{autoesl} and Cynthesizer~\cite{forte}) and is
crucial to producing hardware with high throughput and low latency by allowing temporal overlap of successive loop
iterations. Certifying the correspondence between a sequential design and its pipelined counterpart is challenging due to the huge semantic gap between the two
designs. As a result, hardware designers are vary of using current behavioral synthesis tools as
they are often deemed either (a) aggressively optimized but error-prone or (b) 
reliable but overly conservative, thus often producing circuits of poor quality
or performance~\cite{spark,kundu2008}. Therefore, ensuring
correctness of behaviorally synthesized pipeline designs
is a critical issue in bringing behavioral synthesis into practice.

An approach for certifying loop pipelining transformations using a combination of SEC and 
theorem proving techniques has been proposed earlier~\cite{hrx:dac-12}. The most critical and complex
component of this approach is developing
a loop pipelining algorithm with two key properties: (1) it generates a reference pipeline model 
by exploiting pipeline generation information from the synthesis flow (e.g., the iteration interval 
of a generated pipeline) and the reference model can be compared with a pipelined RTL 
implementation using SEC effectively, and (2) it can be mechanically verified to correctly preserve the semantics of
sequential (non-pipelined) specification of loop execution. The viability of 
this approach was shown by comparing pipeline generated from their algorithm with RTL implementation 
using SEC. However, this algorithm was not certified as well as incomplete. Certification is a key component 
without which correctness of behavioral synthesis process for pipelined designs cannot be claimed.
Our work on developing a certified loop pipelining algorithm using our framework of certified 
primitives is important to facilitating formal verification of behaviorally synthesized
 pipeline designs.

\medskip
The key contributions of this paper are:
\begin{enumerate}
\item Identifying the key provable primitives essential in pipelining algorithms for behavioral synthesis and certifying these primitives in ACL2 theorem prover;
\item Formalizing an invariant to link the sequential loop before pipelining with the pipelined loop;
\item Developing our own executable loop pipelining algorithm in ACL2 using those primitives and certifying this algorithm using ACL2 theorem prover;
\item Testing our certified loop pipelining algorithm on industrial-strength designs
\end{enumerate}

The remainder of this paper is organized as
follows. Section~\ref{sec:background} provides background on the overall project and 
explains the context of our theorem proving work. Section~\ref{sec:formalization} 
discusses our formalization of the intermediate representations used in behavioral synthesis.
We also discuss the correctness statement for loop pipelining algorithms. Section~\ref{sec:challenges} 
discusses the research challenges.
Section~\ref{sec:pipelining-algorithm} discusses our framework and a certified loop pipelining 
algorithm we have developed using the framework. Section~\ref{sec:proof} provides a proof 
sketch for our algorithm. Section~\ref{sec:SEC} provides evaluation of robustness and 
scalability of our algorithm on industrial-strength designs. The related work is discussed in Section~\ref{chapter:related-work}.
We then conclude and discuss future work in Section~\ref{sec:research-plan}.


