1. How is your technique different from what others are doing
2. Under what constraints would this algorithm not work - read his paper carefully
3. Is there no way to prove this algorithm by theorem proving directly
4. Difference bewteen function and loop pipelining
5. Burch and Dill - and manolis - read their papers carefully.
6. difference between ours and software pipelines
7. difference betweeen ours and architectural pipelines
8. read abt phi statements
9. In what shape do we get information from a behavioral synthesis tool.
10. what if a phi-statement prdouces undefined - what do you mean by that.
11.  frame rule, like the one given below, captures the idea that, given a program c with precondition p that holds before it runs and postcondition q that holds afterward, some disjoint condition r should hold both before and after c runs.
12. clock cycle Z
13. hardware resources - shadow reg

form factor
transistor density
how hand-crafted RTL worked initially
equivalence checking


Derek:
---------------------------------------
Ques 1:
pg 11
->correspondence
->granularity ? - fine grained
->orthogonality ? - code evolution

Ques2:
pg 13
code file mark
code anchor
anchor is in both code and implementation, so which way do you go - which anchor you pick first? specification is english, so what do you mean by anchor in the implementation.

Ques3:
AST based anchor
what if you change the name of the function, some misspelling or something then?

Ques 4:
pg 21
why didn't they work? what were the problems? were there common problems?
--------------------------------------------

Zhenkun Yang
-------------
pg 6
view this as a black box - transformation ?

pg 8
should put number on what is happening

-?why would calling N functions in order create path explosion for you?

-> what is they call each other? 

-> pg 14
replace bar(i) with (B, C) = BAR(i, A, C)

-> pg16 why is t4 going along path?

where they are different and make sure other statements remain same?

what abt loop unrolling